\documentclass{article}
\usepackage[left=1.8cm,right=1.8cm,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{enumitem}

\begin{document}
    \hspace{-5mm}\textbf{Harish Rajagopal}\\
    Department of Computer Science and Engineering\\
    Indian Institute of Technology Kanpur, India\\
    \textit{Email: rharish@iitk.ac.in}\\[0.75mm]

    \hrule
    \vspace{1mm}
    \hrule
    \vspace{2mm}
    I am highly interested in Machine Learning. I have done some projects on Computer Vision, Natural Language Processing and Reinforcement Learning with Deep Learning. I am also a metalhead who loves to play guitar.\\[-2.5mm]
    \hrule
    \vspace{1mm}
    \hrule
    \vspace{2mm}

    \textbf{Education}\\[-0.4mm]

    \begin{tabular}{| l | l | l | l |}
        \hline
        Secondary & 2014 & Bharatiya Vidya Bhavan & 93.6\%\\
        \hline
        Sr. Secondary & 2016 & Late P. Jog Junior College & 90.46\%\\
        \hline
        BTech & 2016 & Indian Institute of Technology, Kanpur & 9.74\\
        \hline
    \end{tabular}
    \vspace{5.75mm}

    \textbf{Course details}\\[-0.4mm]

    \begin{tabular}{| l | l | l |}
        \hline
        \textbf{Semester} & \textbf{Total credits} & \textbf{Highlights}\\
        \hline
        Aug 2016 - Dec 2016 & 53 & ESC - 101, PSY - 151\\
        \hline
        Jan 2017 - May 2017& 51 & MTH - 102, TA - 101\\
        \hline
        Aug 2017 - Dec 2017 & 55 & ESC - 201\\
        \hline
    \end{tabular}
    \vspace{8mm}

    \textbf{Skills}\\

    Computer Science \hspace{12.3mm}: \hspace{1mm} Machine learning, Deep Learning, Computer Vision, NLP, Reinforcement Learning\par
    Programming languages \hspace{3mm}: \hspace{1mm} Python, C/C++, Octave, \LaTeX, Bash, HTML\\[+1mm]

    \textbf{Projects}
    \begin{enumerate}[leftmargin=12.5mm]
        \item \textbf{Machine Learning Intern}, \textit{New York Office, IIT Kanpur}
            \begin{enumerate}
                \setlength\itemsep{0em}
                \small
                
                \item Created scrapers for various websites to obtain content using {\it BeautifulSoup} and store scraped data on Couchbase.
                \item Created and trained a {\it Word2Vec} model using {\it Gensim} to compare scraped article contents using {\it Word Mover's Distance}.
                \item Trained a character recognition CNN with sliding windows to scan images to identify English alphanumeric characters.
            \end{enumerate}
        \item \textbf{Depression Therapy Chatbot using Sentiment Analysis}
            \begin{enumerate}
                \setlength\itemsep{0em}
                \small
            
                \item Used a {\it Word2Vec} model to create a {\it Sentiment Analysis} model in Python using LSTMs in Keras.
                \item Implemented a chatbot using manually-created responses dependent on sentiment classification.
                \item Used the {\it Heroku} platform to host the bot and integrated it with Facebook Messenger in Python.
            \end{enumerate}
        \item \textbf{Learn to Play Atari Games using Reinforcement Learning}
            \begin{enumerate}
                \setlength\itemsep{0em}
                \small
                
                \item Used {\it Dynamic Programming} techniques for policy and value iteration to solve an {\it MDP} using {\it OpenAI Gym} environments.
                \item Implemented on-policy and off-policy {\it Monte Carlo} control, {\it SARSA}, {\it Q-Learning} and {\it DQNs} for {\it MDPs} in Python.
                \item Wrote programs to learn playing Atari Pong using both Policy Gradients and DQNs separately.
            \end{enumerate}
    \end{enumerate}

    \textbf{Awards and achievements}
    \begin{itemize}[leftmargin=12.5mm]
        \item[-] All India Rank 185 in JEE Advanced
        \item[-] All India Rank 205 in JEE Mains
        \item[-] A* grade in ESC - 101, PSY - 151, MTH - 102, TA - 101
    \end{itemize}

    \textbf{Hobbies}
    \begin{itemize}[leftmargin=12.5mm]
        \item[-] Music: Electric Guitar, Acoustic Guitar, Bass
        \item[-] Video Gaming: RPGs, Racing, FPS, Sports
        \item[-] Others: Reading novels
    \end{itemize}
\end{document}
